Wheels of Misfortune: Exploring Auto Theft Trends in Ottawa

### 2024-04-23 paper/paper.qmd
new updates

```{r}
#| label: averageTheftYear
#| fig-cap: "Average Auto Theft Crime Reported in ottawa from 2018-2023"
#| echo: false
#| warning: false
#| message: false

@@

```
```{r}
#| label: theft-brands
#| fig-cap: "Number of Cars Stolen by Car Brand in 2018"
#| fig-subcap: ["Illnesses","Visits to the doctor"]
#| layout-ncol: 3
#| echo: false
#| warning: false
#| message: false

@@

brands2018_top10 <- brand_theft_2018_table |>
  filter(n > 26, n != 95) |> 
  mutate(vehicle_make = recode(vehicle_make,
                             "CHEV" = "Chevrolet",
                             "DODG" = "Dodge",
                             "FORD" = "Ford",
                             "GMC" = "GMC",
                             "HOND" = "Honda",
                             "HYUN" = "Hyundai",
                             "LEXS" = "Lexus",
                             "MAZD" = "Mazda",
                             "NISS" = "Nissan",
                             "TOYT" = "Toyota"
                             ))
brands2018_top10 |>
  ggplot(aes(x = vehicle_make, y = n, fill = n)) +  # Use n as the fill aesthetic
  geom_bar(stat = "identity") +
  theme_minimal() + theme(legend.position = "bottom") + 
  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2018")

@@

# filter for Top 10 Brands 
brandtheft2019_top10 <- brand_theft_2019 |>
                          count(vehicle_make) |>
                          filter(n > 25, n != 118) |> 
                          mutate(vehicle_make = recode(vehicle_make,
                                                        "CHEV" = "Chevrolet",
                                                        "DODG" = "Dodge",
                                                        "FORD" = "Ford",
                                                        "GMC" = "GMC",
                                                        "HOND" = "Honda",
                                                        "HYUN" = "Hyundai",
                                                        "LEXS" = "Lexus",
                                                        "MAZD" = "Mazda",
                                                        "NISS" = "Nissan",
                                                        "TOYT" = "Toyota"
                                                     ))
brandtheft2019_top10 |>
  ggplot(aes(x = vehicle_make, y = n, fill = n)) +  # Use n as the fill aesthetic
  geom_bar(stat = "identity") +
  theme_minimal() + theme(legend.position = "bottom") + 
  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2019")

@@

# filter for Top 10 Brands 
brandtheft2020_top10 <- brand_theft_2020 |>
                          count(vehicle_make) |>
                          filter(n > 21, n != 95) |> 
                          mutate(vehicle_make = recode(vehicle_make,
                                                        "CHEV" = "Chevrolet",
                                                        "DODG" = "Dodge",
                                                        "FORD" = "Ford",
                                                        "GMC" = "GMC",
                                                        "HOND" = "Honda",
                                                        "HYUN" = "Hyundai",
                                                        "LEXS" = "Lexus",
                                                        "MERZ" = "Mercedes-Benz ",
                                                        "MAZD" = "Mazda",
                                                        "NISS" = "Nissan",
                                                        "TOYT" = "Toyota"
                                                     ))
brandtheft2020_top10 |>
  ggplot(aes(x = vehicle_make, y = n, fill = n)) +  # Use n as the fill aesthetic
  geom_bar(stat = "identity") +
  theme_minimal() + theme(legend.position = "bottom") + 
  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2020")

@@

# filter for Top 10 Brands 
brandtheft2021_top10 <- brand_theft_2021 |>
                          count(vehicle_make) |>
                          filter(n > 21, n != 115) |> 
                          mutate(vehicle_make = recode(vehicle_make,
                                                        "CHEV" = "Chevrolet",
                                                        "DODG" = "Dodge",
                                                        "FORD" = "Ford",
                                                        "GMC" = "GMC",
                                                        "HOND" = "Honda",
                                                        "HYUN" = "Hyundai",
                                                        "LEXS" = "Lexus",
                                                        "MERZ" = "Mercedes-Benz ",
                                                        "MAZD" = "Mazda",
                                                        "NISS" = "Nissan",
                                                        "TOYT" = "Toyota"
                                                     ))
brandtheft2021_top10 |>
  ggplot(aes(x = vehicle_make, y = n, fill = n)) +  # Use n as the fill aesthetic
  geom_bar(stat = "identity") +
  theme_minimal() + theme(legend.position = "bottom") + 
  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2021")

@@

# filter for Top 10 Brands 
brandtheft2022_top10 <- brand_theft_2022 |>
                          count(vehicle_make) |>
                          filter(n > 34, n != 124) |> 
                          mutate(vehicle_make = recode(vehicle_make,
                                                        "CHEV" = "Chevrolet",
                                                        "DODG" = "Dodge",
                                                        "FORD" = "Ford",
                                                        "GMC" = "GMC",
                                                        "HOND" = "Honda",
                                                        "HYUN" = "Hyundai",
                                                        "LEXS" = "Lexus",
                                                        "MERZ" = "Mercedes-Benz ",
                                                        "MAZD" = "Mazda",
                                                        "NISS" = "Nissan",
                                                        "TOYT" = "Toyota"
                                                     ))
brandtheft2022_top10 |>
  ggplot(aes(x = vehicle_make, y = n, fill = n)) +  # Use n as the fill aesthetic
  geom_bar(stat = "identity") +
  theme_minimal() + theme(legend.position = "bottom") + 
  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2022")

@@

# filter for Top 10 Brands 
brandtheft2023_top10 <- brand_theft_2023 |>
                          count(vehicle_make) |>
                          filter(n > 16, n != 36) |> 
                          mutate(vehicle_make = recode(vehicle_make,
                                                       "ACUR" = "Acura",
                                                        "CHEV" = "Chevrolet",
                                                        "DODG" = "Dodge",
                                                        "FORD" = "Ford",
                                                        "GMC" = "GMC",
                                                        "HOND" = "Honda",
                                                        "HYUN" = "Hyundai",
                                                        "LEXS" = "Lexus",
                                                        "MERZ" = "Mercedes-Benz ",
                                                        "MAZD" = "Mazda",
                                                        "NISS" = "Nissan",
                                                        "TOYT" = "Toyota"
                                                     ))
brandtheft2023_top10 |>
  ggplot(aes(x = vehicle_make, y = n, fill = n)) +  # Use n as the fill aesthetic
  geom_bar(stat = "identity") +
  theme_minimal() + theme(legend.position = "bottom") + 
  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2023")

@@

```{r}
#| label: theft-recovery
#| fig-cap: "Recovery Rates from 2018 - 2023"
#| echo: false
#| warning: false
#| message: false

theft_recovery <- auto_Data |>
  select(theft_year, recovered)

theft_recovery |>
  ggplot(mapping = aes(x = theft_year, fill = recovered)) +
  geom_bar(position = "dodge2") +
  labs(x = "Age group", y = "Number of observations", fill = "Vote") +
  theme(legend.position = "bottom")
```
```{r}
#| label: theft-neighborhoods
#| fig-cap: "Recovery Rates from 2018 - 2023"
#| echo: false
#| warning: false
#| message: false

theft_recovery <- auto_Data |>
  select(neighbourhood) |> 
  count(neighbourhood) |> 
  filter(n > 180)

theft_recovery |>
  ggplot(aes(x = neighbourhood, y = n, fill = n)) +  # Use n as the fill aesthetic
  geom_bar(stat = "identity") +
  theme_minimal() + theme(legend.position = "bottom") + 
  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2022") 
```

>>>>

```{r}
#| label: averageTheftYear
#| fig-cap: "Average Auto Theft Crimes Reported in ottawa from 2018-2023"
#| echo: false
#| warning: false
#| message: false

@@

```
```{r}
#| label: theftBrands
#| fig-cap: "Top 10 Car Brands Stolen from 2018-2023"
#| fig-subcap: ["Top 10 Brands Affected in 2018", "Top 10 Brands Affected in 2019", "Top 10 Brands Affected in 2020", "Top 10 Brands Affected in 2021", "Top 10 Brands Affected in 2022", "Top 10 Brands Affected in 2023"]
#| layout-ncol: 2
#| echo: false
#| warning: false
#| message: false

@@

brands2018_top10 <- brand_theft_2018_table |>
  filter(n > 26, n != 95) |> 
  mutate(vehicle_make = recode(vehicle_make,
                             "CHEV" = "Chevrolet",
                             "DODG" = "Dodge",
                             "FORD" = "Ford",
                             "GMC" = "GMC",
                             "HOND" = "Honda",
                             "HYUN" = "Hyundai",
                             "LEXS" = "Lexus",
                             "MAZD" = "Mazda",
                             "NISS" = "Nissan",
                             "TOYT" = "Toyota"
                             ))
brands2018_top10 |>
  ggplot(aes(x = vehicle_make, y = n, fill = n)) +  # Use n as the fill aesthetic
  geom_bar(stat = "identity") +
  theme_minimal() + theme(legend.position = "bottom") + 
  labs(x = "Vehicle Brand", y = "Count", fill = "Count")

@@

# filter for Top 10 Brands 
brandtheft2019_top10 <- brand_theft_2019 |>
                          count(vehicle_make) |>
                          filter(n > 25, n != 118) |> 
                          mutate(vehicle_make = recode(vehicle_make,
                                                        "CHEV" = "Chevrolet",
                                                        "DODG" = "Dodge",
                                                        "FORD" = "Ford",
                                                        "GMC" = "GMC",
                                                        "HOND" = "Honda",
                                                        "HYUN" = "Hyundai",
                                                        "LEXS" = "Lexus",
                                                        "MAZD" = "Mazda",
                                                        "NISS" = "Nissan",
                                                        "TOYT" = "Toyota"
                                                     ))
brandtheft2019_top10 |>
  ggplot(aes(x = vehicle_make, y = n, fill = n)) +  # Use n as the fill aesthetic
  geom_bar(stat = "identity") +
  theme_minimal() + theme(legend.position = "bottom") + 
  labs(x = "Vehicle Brand", y = "Count", fill = "Count")

@@

# filter for Top 10 Brands 
brandtheft2020_top10 <- brand_theft_2020 |>
                          count(vehicle_make) |>
                          filter(n > 21, n != 95) |> 
                          mutate(vehicle_make = recode(vehicle_make,
                                                        "CHEV" = "Chevrolet",
                                                        "DODG" = "Dodge",
                                                        "FORD" = "Ford",
                                                        "GMC" = "GMC",
                                                        "HOND" = "Honda",
                                                        "HYUN" = "Hyundai",
                                                        "LEXS" = "Lexus",
                                                        "MERZ" = "Mercedes",
                                                        "MAZD" = "Mazda",
                                                        "NISS" = "Nissan",
                                                        "TOYT" = "Toyota"
                                                     ))
brandtheft2020_top10 |>
  ggplot(aes(x = vehicle_make, y = n, fill = n)) +  # Use n as the fill aesthetic
  geom_bar(stat = "identity") +
  theme_minimal() + theme(legend.position = "bottom") + 
  labs(x = "Vehicle Brand", y = "Count", fill = "Count")

@@

# filter for Top 10 Brands 
brandtheft2021_top10 <- brand_theft_2021 |>
                          count(vehicle_make) |>
                          filter(n > 21, n != 115) |> 
                          mutate(vehicle_make = recode(vehicle_make,
                                                        "CHEV" = "Chevrolet",
                                                        "DODG" = "Dodge",
                                                        "FORD" = "Ford",
                                                        "GMC" = "GMC",
                                                        "HOND" = "Honda",
                                                        "HYUN" = "Hyundai",
                                                        "LEXS" = "Lexus",
                                                        "MERZ" = "Mercedes-Benz ",
                                                        "MAZD" = "Mazda",
                                                        "NISS" = "Nissan",
                                                        "TOYT" = "Toyota"
                                                     ))
brandtheft2021_top10 |>
  ggplot(aes(x = vehicle_make, y = n, fill = n)) +  # Use n as the fill aesthetic
  geom_bar(stat = "identity") +
  theme_minimal() + theme(legend.position = "bottom") + 
  labs(x = "Vehicle Brand", y = "Count", fill = "Count")

@@

# filter for Top 10 Brands 
brandtheft2022_top10 <- brand_theft_2022 |>
                          count(vehicle_make) |>
                          filter(n > 34, n != 124) |> 
                          mutate(vehicle_make = recode(vehicle_make,
                                                        "CHEV" = "Chevrolet",
                                                        "DODG" = "Dodge",
                                                        "FORD" = "Ford",
                                                        "GMC" = "GMC",
                                                        "HOND" = "Honda",
                                                        "HYUN" = "Hyundai",
                                                        "LEXS" = "Lexus",
                                                        "MERZ" = "Mercedes-Benz ",
                                                        "MAZD" = "Mazda",
                                                        "NISS" = "Nissan",
                                                        "TOYT" = "Toyota"
                                                     ))
brandtheft2022_top10 |>
  ggplot(aes(x = vehicle_make, y = n, fill = n)) +  # Use n as the fill aesthetic
  geom_bar(stat = "identity") +
  theme_minimal() + theme(legend.position = "bottom") + 
  labs(x = "Vehicle Brand", y = "Count", fill = "Count")

@@

# filter for Top 10 Brands 
brandtheft2023_top10 <- brand_theft_2023 |>
                          count(vehicle_make) |>
                          filter(n > 16, n != 36) |> 
                          mutate(vehicle_make = recode(vehicle_make,
                                                       "ACUR" = "Acura",
                                                        "CHEV" = "Chevrolet",
                                                        "DODG" = "Dodge",
                                                        "FORD" = "Ford",
                                                        "GMC" = "GMC",
                                                        "HOND" = "Honda",
                                                        "HYUN" = "Hyundai",
                                                        "LEXS" = "Lexus",
                                                        "MERZ" = "Mercedes-Benz ",
                                                        "MAZD" = "Mazda",
                                                        "NISS" = "Nissan",
                                                        "TOYT" = "Toyota"
                                                     ))
brandtheft2023_top10 |>
  ggplot(aes(x = vehicle_make, y = n, fill = n)) +  # Use n as the fill aesthetic
  geom_bar(stat = "identity") +
  theme_minimal() + theme(legend.position = "bottom") + 
  labs(x = "Vehicle Brand", y = "Count", fill = "Count")

@@

```{r}
#| label: theft-recovery
#| fig-cap: "Recovery Rates from 2018-2023"
#| echo: false
#| warning: false
#| message: false

# Filter for Count of Cars Recovered from 2018-2023
theft_recovery <- auto_Data |>
  select(theft_year, recovered)

# Create a Stacked Bar Graph of the Data
theft_recovery |>
  ggplot(mapping = aes(x = theft_year, fill = recovered)) +
  geom_bar(position = "dodge2") +
  labs(x = "Age group", y = "Number of observations", fill = "Vote") +
  theme(legend.position = "bottom")

```{r}
#| label: theft-neighborhoods
#| fig-cap: "Top 5 Neighborhoods with Most Auto Thefts from 2018-2023"
#| echo: false
#| warning: false
#| message: false

# Filter for top 10 neighborhoods with the highest auto theft rates from 2018-2023
theft_recovery <- auto_Data |>
  select(neighbourhood) |> 
  count(neighbourhood) |> 
  filter(n > 180)

theft_recovery |>
  ggplot(aes(x = neighbourhood, y = n, fill = n)) +  # Use n as the fill aesthetic
  geom_bar(stat = "identity") +
  theme_minimal() + theme(legend.position = "bottom") + 
  labs(x = "Vehicle Brand", y = "Count", fill = "Count")

```{r}
#| label: vehicleYearStolen
#| fig-cap: "Vehicle Years Stolen from 2018-2023"
#| echo: false
#| warning: false
#| message: false

# Count for vehicle year 
years_vehicle <- auto_Data |>
  select(vehicle_year) |> 
  count(vehicle_year)

# Code from Chat GPT (see LLM)

years_vehicle <- years_vehicle |>
  mutate(vehicle_year = as.numeric(vehicle_year))  # Convert to numeric if 'vehicle_year' is factor

# Create a scatter plot
ggplot(years_vehicle, aes(x = vehicle_year, y = n, fill = n)) +
  geom_point() +  # Use geom_point() for scatter plot
  scale_x_continuous(limits = c(1930, 2024)) +  # Adjust the x-axis limits here
  theme_minimal() + 
  theme(legend.position = "bottom") + 
  labs(x = "Vehicle Brand", y = "Count", fill = "Count")
